---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/index-",
  out.width = "100%"
)
```

# disputeR

[![R build status](https://github.com/chris-prener/disputeR/workflows/R-CMD-check/badge.svg)](https://github.com/chris-prener/disputeR/actions)

The goal of `disputeR` is to easily implement unit tests within functions and Shiny modules. Unlike other packages that provide frameworks for developing these tests, `disputeR` takes a different approach, offering pre-built unit tests that can easily be dropped into functions and test for basic requirements, like class, length, and common values used in function writing (such as `NULL`). These tests return informative error messages that follow the [`tidyverse` style guide's recommendations](https://style.tidyverse.org/errors.html).

## Basic Usage

The following example creates a simple function for squaring a numeric value, which is passed to the function using the parameter `x`. Two `disputeR` functions are added before squaring `x`. The first, `dis_not_missing()`, evaluates `x` to ensure the parameter is not skipped by the user. The second, `dis_numeric()`, applies a number of checks to `x`.

```{r}
## load package
library(disputeR)

## define example function
example <- function(x){

   ## check inputs with disputeR
   dis_not_missing(.f = missing(x))
   dis_numeric(x, null_valid = FALSE)

   ## square
   out <- x^2

   ## return output
   return(out)

}
```

The checks `dis_numeric()` applies include default settings that require a scalar input (i.e. `length(x) == 1`) and do not allow `NA`, `NaN`, `Inf`, or `-Inf`. Each of these can be enabled using the parameters for `dis_numeric()` at the function author's discretion. In addition, we enforce a requirement that `x` cannot be `NULL`.

In the case of the following example, `2` is a valid input to `example()` and the function executes without error:

```{r}
example(x = 2)
```

However, if the user does not specify the `x` argument, an informative error message is returned along with instructions for how to rectify the issue:

```{r, error = TRUE}
example()
```

Likewise, if `x` is incorrectly specified (such as with a `class(x) == character` value), a similar message along with instructions for fixing the error is returned:

```{r, error = TRUE}
example(x = "test")
```

## Additional Resources
Check out the vignettes and reference page for additional resources on using `disputeR` in your own projects as well as developing packages that contain `disputeR` checks.

## Code of Conduct
  
Please note that the disputeR project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
