---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# disputeR

[![R build status](https://github.com/chris-prener/disputeR/workflows/R-CMD-check/badge.svg)](https://github.com/chris-prener/disputeR/actions)

The goal of `disputeR` is to easily implement unit tests within functions and Shiny modules. Unlike other packages that provide frameworks for developing these tests, `disputeR` takes a different approach, offering pre-built unit tests that can easily be dropped into functions and test for basic requirements, like class, length, and common values used in function writing (such as `NULL`). These tests return informative error messages that follow the 'tidyverse' style guide's recommendations.

## Installation

You can install the development version of `disputeR` from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("chris-prener/disputeR")
```

## Basic Usage

The following example creates a simple function for squaring a numeric value, which is passed to the function using the parameter `x`. Two `disputeR` functions are added before squaring `x`. The first, `dis_not_missing()`, evaluates `x` to ensure the parameter is not skipped by the user. The second, `dis_numeric()`, applies a number of checks to `x`.

```{r}
## load package
library(disputeR)

## define example function
example <- function(x){

   ## check inputs with disputeR
   dis_not_missing(.f = missing(x))
   dis_numeric(x, null_valid = FALSE)

   ## square
   out <- x^2

   ## return output
   return(out)

}
```

The checks `dis_numeric()` applies include default settings that require a scalar input (i.e. `length(x) == 1`) and do not allow `NA`, `NaN`, `Inf`, or `-Inf`. Each of these can be enabled using the parameters for `dis_numeric()` at the function author's discretion. In addition, we enforce a requirement that `x` cannot be `NULL`.

In the case of the following example, `2` is a valid input to `example()` and the function executes without error:

```{r}
example(x = 2)
```

However, if the user does not specify the `x` argument, an informative error message is returned along with instructions for how to rectify the issue:

```{r, error = TRUE}
example()
```

Likewise, if `x` is incorrectly specified (such as with a `class(x) == character` value), a similar message along with instructions for fixing the error is returned:

```{r, error = TRUE}
example(x = "test")
```

A similar approach can be taken for both character and logical arguments. For example :

```{r}
example <- function(x, round = FALSE){

   ## check inputs with disputeR
   dis_not_missing(.f = missing(x))
   dis_numeric(x, null_valid = FALSE)
   dis_logical(round, null_valid = FALSE)

   ## square
   out <- x^2
   
   ## optionally round
   if (isTRUE(round)){
     out <- round(out, digits = 2)
   }

   ## return output
   return(out)

}
```

The use of `dis_logical()` ensures that a non-`NULL`, logical argument (either `TRUE` or `FALSE`) is passed to the `round` parameter. This allows us to validate or dispute user inputs:

```{r, error = TRUE}
## no issues
example(x = 2.25)
example(x = 2.25, round = TRUE)

## error found
example(x = 2.25, round = "yes")
```

Both `dis_numeric()` and `dis_character()` also accept strings of valid values. The following function allows for only certain words to be passed to the `x` parameter:

```{r}
example <- function(x){

   ## check inputs with disputeR
   dis_not_missing(.f = missing(x))
   dis_character(x, valid = c("ham", "eggs", "bacon"), null_valid = FALSE,)

   ## create output
   out <- paste0(x, " is delicious!")

   ## return output
   return(out)

}
```

The value for `x` will be checked against the vector passed to the `valid` parameter in addition to the other checks that `dis_character()` performs:

```{r, error = TRUE}
## no issues
example(x = "bacon")

## error found
example(x = "spam")
```

The `valid` argument works the same way with `dis_numeric()`, passing either integer or numeric values to `valid` instead of character data.

## Developing with disputeR

There are two functions that are designed for modifying the developer experience. The first is `dis_fact_check()`. If you are working on a Shiny app in particular and include `disputeR` functions in your modules or functions, setting this function to `FALSE` at the beginning of `global.R` or `app.R` will stop the core functions (`dis_character()`, `dis_logical()`, `dis_not_missing()`, and `dis_numeric()`) from executing the vast majority of their checks. 

```r
# beginning of global.R or app.R
dis_fact_check(fact_check = FALSE)
```

This is useful in situations where you have completed development of the app and have all functions specified correctly. In that instance, `disputeR` does not add much to your app except milliseconds of execution time. Turning fact checking off saves most of those milliseconds for your end users. This can be overridden on a function call by function call basis with the `fact_check` argument, allowing you to selectively use `disputeR` on user inputs while disabling it elsewhere in your app.

If you run into issues with one of your functions or modules, simply comment out `dis_fact_check()` and restart your `R` session to get the benefits of `disputeR` back.

The `dis_dev_check()` function is also designed to save your end users milliseconds of execution time when `disputeR` and is particularly aimed at package developers. Since end users will not be interacting with `disputeR` directly, the logic checks on `disputeR`'s own functions are disabled by default. This is done to save your end users execution time, but comes at the expense of the package developer's experience. Using `dis_dev_check()` and setting it to `TRUE` will allow `disptueR` to validate its own parameters, easing your own use of the package as a developer.

```r
# beginning of development session
dis_dev_check(dev_check = TRUE)
```

## Code of Conduct
  
Please note that the disputeR project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
